A Comparative Quality Metric for Untargeted Fuzzing with Logic State Coverage

### 2024-09-18 paper/main.tex
Draft for arXiv

%To satisfy the requirements, we incorporate a \emph{bloom filter} to manage a
%set of observed logic states. 
In this section, we first present the design overview (\autoref{s:design:over})
and elaborate on details (\autoref{s:design:rec} and \ref{s:design:count}).

@@

@@

@@

@@

\begin{itemize}
  \item \textbf{RQ1} Can logic state coverage be implemented with all
    requirements in \autoref{s:measure:req} satisfied?
  \item \textbf{RQ2} How does logic state coverage measure the comparative
    quality of popular untargeted fuzzers?
  \item \textbf{RQ3} Can logic state coverage distinguish the quality of the
    fuzzers indistinguishable from conventional edge coverage?
\end{itemize}

@@

One possible compromise is performing preliminary static analysis to
\emph{estimate} the total number of program behaviors, including abnormal ones.
However, it would pose a potential problem that the estimated 100\% completeness
does not necessarily mean the actual 100\% completeness, which may compromise
the very purpose of measuring absolute completeness.

@@

>>>>

%To satisfy the requirements, we incorporate a \emph{bloom filter} to manage a
%set of observed logic states. 
In this section, we first present the design overview (\autoref{s:design:over})
and elaborate on details (\autoref{s:design:rec} and
\ref{s:design:count}). The implementation in progress can be found at
\url{https://github.com/gwangmu/LogicStateCoverage}.

@@

Recording a logic state is equivalent to tracing executed edges in conventional
edge coverage. Specifically, a taken edge is calculated by combining the
hashes of two consecutively executed basic blocks, which is then added to the
logic state of the current execution. Notice that a logic state is \textit{per
execution} and is not merged throughout multiple executions as in edge
coverage. We implemented logic state recoding based on the LLVM pass for
\cc{afl-clang-fast} \cite{afl}.

@@

The recorded logic state is updated to the bloom filter at the end of every
execution. In particular, a number of hashes are calculated from the entire
logic state (i.e., \cc{A}, \cc{B}, \cc{C}, and \cc{D} in \autoref{f:design}) and
are used as bloom filter indices to be set. We used MurMurHash3
\cite{murmurhash3} to hash a logic state.

@@

Finally, the logic state coverage of the fuzzing campaign is measured by
periodically reading off the bloom filter, where the number of 1s in the filter
is fed to \autoref{e:count} to count the distinctive number of observed logic
states. We implemented logic state coverage measurement as a separate process,
which a measurer should launch before starting a fuzzing campaign.

@@

\begin{itemize}
  \item \textbf{RQ1} Can logic state coverage be implemented with all
    requirements in \autoref{s:measure:req} satisfied?
  \item \textbf{RQ2} How does logic state coverage measure the comparative
    quality of popular untargeted fuzzers?
  \item \textbf{RQ3} Can logic state coverage distinguish the fuzzers
    indistinguishable from conventional edge coverage?
\end{itemize}

@@

One possible compromise is performing preliminary static analysis to
\emph{estimate} the total number of program behaviors, including abnormal ones.
However, it would pose a potential problem that the estimated 100\% completeness
does not necessarily mean the actual 100\% completeness, defeating the very
purpose of measuring \textit{completeness}.

@@

While fuzzing is widely accepted as an efficient program testing technique, 
it still lacks the metric to comparatively measure the quality of various
designs and implementations. To establish a reliable quality metric, we
note that the more interesting behaviors a fuzzer has observed, the stronger
guarantee it can provide about the absence of unknown abnormal behaviors. Based
on this, we propose \emph{logic state coverage} as a proxy metric to count
observed interesting behaviors. A logic state is a set of satisfied branches
during one execution, where its coverage is the count of \emph{individual}
observed logic states during a fuzzing campaign. We implemented logic state
coverage using a bloom filter and performed a preliminary evaluation with AFL++
and XMLLint.
